%\chapter{Technologies}
\section{Interesting facts about Sombrero}
    Here is a List of some interesting facts about sombrero:
    \begin{itemize}
        \item \textbf{Amount of Scala files:} 52
        \item \textbf{Amount of device widgets:} 5
        \item \textbf{Why we chose Scala:}
            Scala is lightweight and mostly boilerplate free.
        \item \textbf{Why we chose Lift:}

            David Pollak on Lift:

        \begin{quote}
            ``Lift is one of the most secure web frameworks around because, by default, it's resistant to replay attacks, cross site scripting, cross site request forgeries, etc.

            Lift's comet support is the best around.

            Lift's Ajax support is easier than any other web framework that I know of.

            Comparison to other Frameworks:
            \begin{itemize}
                \item \textbf{Comparison to Grails}

                    Lift is faster than Grails (Groovy is slow).
                \item \textbf{Usability}

                    Lift apps are going to be smaller and more maintainable than either Grails or Spring MVC and because Scala is strongly typed, the compiler will help you.
                \item \textbf{Architecture}

                    Keeping stuff around in XML rather than Strings and associating HTML elements with functions I consider to be vastly architecturally superior to what Grails and Spring MVC do.
                \item \textbf{Deployment}

                    No difference.  A WAR file is a WAR file.
                \item \textbf{Weaknesses of Lift}

                    Where Lift doesn't do better than other web frameworks is generally CRUD related.  Doing CRUD apps in Lift is not the simplest thing in the world. But then, Lift is not tied to a single persistence mechanism (like Grails), so you trade on thing for another.''\cite{google.com:liftgroup}
            \end{itemize}
        \end{quote}
        \item \textbf{Library's:}
        \begin{itemize}
            \item \textbf{JQuery}           1.4.2
            \item \textbf{JQuery UI}        1.7.2
            \item \textbf{Yaml}             3.2.1
            \item \textbf{JQuery-message}   1.0
            \item \textbf{JQuery-collide}   1.0
            \item \textbf{Fancybox}         1.3.0
            \item \textbf{Easing}           1.3
            \item \textbf{Mousewheel}       3.0.2
            \item \textbf{Calimero}         2.0a4
            \item \textbf{Scala Library}    2.7.7
            \item \textbf{JDK}              1.6.0
            \item \textbf{Lift}             1.0
        \end{itemize}
    \end{itemize}
